Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-10 prop_letter.tex
prop_letter.tex Finish Updating the easy parts of Andy's note...

We can now proceed with our technical result which is that the
original Kochen-Specker theorem is a statement regarding the
non-existence of $0$-deterministic QIVPM.

@@

\noindent To explain why this result is equivalent to the original
Kochen-Specker theorem and to prove it at the same time, we proceed by
assuming a 0-deterministic QIVPM $\bar{\mu}$ and derive a
contradiction to the original Kochen-Specker theorem. Instead of
adapting the more complicated proof for $d=3$, our counterexample uses
the simpler proof for a Hilbert space of dimension $d=4$ and is
constructed as follows.

@@

The proof is by contradiction: suppose there is a
$\delta$-deterministic
QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, we claim that the
following
map~$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
  \imposs,\necess\right\} $ is a 0-deterministic QIVPM:
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, since we know that 0-deterministic QIVPMs do not exist, the
map $\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ does
not exist.

@@

Given a three-dimensional Hilbert space $\Hilb$ with an orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
$\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a QIVPM. Then we
will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an empty set
by contradiction. Suppose there is a state
$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since
$\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$ for any~$P$, the
state~$\rho$ must satisfy:
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0
\end{equation}
which implies
$\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$ for all
$j$. However, the last equation never holds.  Thus, there is no state
consistent with the QIVPM on all projectors~$P\in\events$, so
Gleason's theorem does not hold for at least this example.

@@

We call a state $\rho$ a \emph{totally mixed state} if
$0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any
$P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $. We also
define a notion of distance based on the following norm of a
matrix~$A$ as
$\left\Vert A\right\Vert
=\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}. Now
consider the following collection of intervals:
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{ \imposs, \necess \}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at each end.

>>>>

We can now proceed with our technical result, which is that the
original Kochen-Specker theorem is a statement regarding the
non-existence of $0$-deterministic QIVPM.

@@

\noindent To explain why this result is equivalent to the original
Kochen-Specker theorem and to prove it at the same time, we proceed by
assuming a 0-deterministic QIVPM $\bar{\mu}$ and derive the same
contradiction as the original Kochen-Specker theorem. Instead of
adapting the more complicated proof for $d=3$, our counterexample uses
the simpler proof for a Hilbert space of dimension $d=4$ and is
constructed as follows.

@@

The proof is by contradiction: Suppose there is a
$\delta$-deterministic
QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$. We claim that the
following
map~$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
  \imposs,\necess\right\} $ is a 0-deterministic QIVPM:
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, since we know that 0-deterministic QIVPMs do not exist, the
map $\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ does
not exist.

@@

Given a three-dimensional Hilbert space $\Hilb$ with an orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
$\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a QIVPM. Then we
will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an empty set
by contradiction. Suppose there is a state
$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since
$\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$ for any~$P$, the
state~$\rho$ must satisfy:
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0
\end{equation}
which implies
$\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$ for all
$j$. However, the last equation never holds.  Thus, there is no state
consistent with the QIVPM on all projectors~$P\in\events$, so
Gleason's theorem does not hold, at least for this example.

@@

We call a state $\rho$ a \emph{totally mixed state} if
$0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any
$P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $. We also
define a notion of distance based on the following norm of a
matrix~$A$ as
$\left\Vert A\right\Vert
=\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}. Now
consider the following collection of intervals:
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{ \imposs, \necess \}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end.
